Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-18 manuscript/letter.tex
Review of the part about gravity modeling
Renames the section to just "Gravity modeling" because that is what it's
about.
Adds an introduction to the section and removes a lot of repetition to
make the text flow better.
Also change some of the language to avoid blaming people too much.

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their objectives and geophysical interpretations are significantly
different.
The corrections applied when calculating the Bouguer anomaly approximate
gravity at the geoid from measurements made on the Earth's surface by removing
the effects of the distance and the topographic masses between the two points.
Apart from the centrifugal effect mentioned in the previous section,
the gravitational effect of topographic masses between the geoid and ellipsoid
are still present in the anomaly.
Conversely, when calculating the Bouguer disturbance one removes from
the observed gravity the effects of an ellipsoidal reference model (the normal
Earth) and the known topographic masses between the surface of the ellipsoid
and observation point.
Hence, the Bouguer disturbance can be interpreted as being solely the
gravitational effect produced by the density contrasts between the actual
internal density distribution of the Earth and that of the normal Earth.

@@

\section{Mathematical description of the gravity disturbance at the TCS}

From eq. \ref{eq:gravity-disturbance-vector}, the gravity vector $\vector{g}_{i}$, at a point
$(x_{i}, y_{i}, z_{i})$ in the TCS (Fig. \ref{fig:TCS}a), can be represented by:
\begin{equation}
\vector{g}_{i} = \vector{\gamma}_{i} + \delta \vector{g}_{i} \: ,
\label{eq:gravity-vector-TCS}
\end{equation}
where $\vector{\gamma}_{i}$ and $\delta \vector{g}_{i}$
are, respectively, the normal gravity vector and the
gravity disturbance vector produced by the anomalous
masses at the point $(x_{i}, y_{i}, z_{i})$.

The gravity $g_{i}$ can be approximated by a first order Taylor's
expansion as follows \citep{sanso_sideris2013}:
\begin{equation}
g_{i} \approx \gamma_{i} +
\versor{\gamma}_{i}^{\top} \delta \vector{g}_{i} \: ,
\label{eq:gobs-approx}
\end{equation}
where $^{\top}$ denotes transposition,
$\versor{\gamma}_{i} = -\versor{u}_{i}$ is a unit
vector with the same direction as the normal gravity vector
$\vector{\gamma}_{i}$ at the point $(x_{i}, y_{i}, z_{i})$, in the
topocentric Cartesian coordinate system (Fig. \ref{fig:TCS}a), and
$\versor{u}_{i}$ is the unit vector $\versor{u}$
(eq. \ref{eq:unit-vectors}) evaluated at the corresponding
point $(h_{i}, \varphi_{i}, \lambda_{i})$ in the
GGS (Fig. \ref{fig:GCS-GGS}).
This approximation can be made because, fortunately,
the condition $\gamma_{i} \gg \| \delta \mathbf{g}_{i} \|$
is met at all points located above or on the Earth's surface.
This approximation is known in geodesy \citep[e.g.,][]{sanso_sideris2013}
and a similar approximation is largely used in applied geophysics
for representing total-field anomalies \citep[e.g.,][]{blakely1996}.
Notice that, in local- or regional-gravity studies, the unit
vector $\versor{\gamma}_{i}$ (eq. \ref{eq:gobs-approx})
may be considered constant throughout the study area and
parallel to the $z$ axis of the TCS (Fig. \ref{fig:TCS}a).
By using the approximation defined in eq.
\ref{eq:gobs-approx}, the gravity disturbance
(eq. \ref{eq:gravity-disturbance}) can be rewritten as
follows
\begin{equation}
\delta g_{i} \approx \versor{\gamma}_{P}^{\top} \delta \vector{g}_{i} \: ,
\label{eq:gravity-disturbance-approx-TCS}
\end{equation}
where $\versor{\gamma}_{P}$ represents the unit vector with
the same direction as the normal gravity vector $\vector{\gamma}_{P}$
at the origin $P$ of the TCS (Fig. \ref{fig:TCS}a).
This equation shows that the gravity disturbance $\delta g_{i}$ (eq. \ref{eq:gravity-disturbance}) is different from the magnitude
of the gravity disturbance vector $\delta \vector{g}_{i}$
produced by the gravity sources. Rather, it represents the
component of $\delta \vector{g}_{i}$ on the direction of the normal
gravity vector \citep{hofmann-wellenhof-moritz2005, sanso_sideris2013}.

In the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance $\delta g_{i}$ (eq. \ref{eq:gravity-disturbance-approx-TCS})
can be defined as the $z$-component of the gravitational attraction exerted by
the gravity sources at the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance produced by
a gravity source can be represented by the following harmonic function:
\begin{equation}
d_{i} = \int\int\limits_{v}\int
\frac{c_{g} \, G \, \Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})
\, (z_{i} - z^{\prime}) \: dv^{\prime}}
{\sqrt{(x_{i} - x^{\prime})^{2} +
(y_{i} - y^{\prime})^{2} + (z_{i} - z^{\prime})^{2}}} \: ,
\label{eq:gz-local}
\end{equation}
where $G$ is the Newtonian constant of gravitation (in $m^{3} \, kg^{-1} \, s^{-2}$),
$c_{g} = 10^{5}$ is a constant factor transforming from $m \, s^{-2}$
to milligal (mGal), $\Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})$
is the density contrast (in $kg \, m^{-3}$) at a point
$(x^{\prime}, y^{\prime}, z^{\prime})$ within
the volume $v$ of the gravity source and the integration is conducted
over $x^{\prime}$, $y^{\prime}$ and $z^{\prime}$.
This equation can be easily generalized for the case
of multiple gravity sources.
Practically, all articles accomplishing gravity modelling
use the quantity $d_{i}$ (eq. \ref{eq:gz-local})
to represent the gravity anomaly produced by
gravity sources \citep[e.g.,][]{blakely1996}.
Consequently, almost all geophysicists use the gravity
anomaly as an approximation of the gravity disturbance
produced by the gravity sources. Notice that $d_{i}$
(eq. \ref{eq:gz-local}) does not depend on the
height with respect to the Geoid (orthometric height).
Rather, it depends on the relative position
of the observation points $(x_{i}, y_{i}, z_{i})$ with respect to
the gravity sources in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to notice that, contrary to what
is written in some basic texts, gravimeters do not measure
the quantity $d_{i}$ (eq. \ref{eq:gz-local}).

>>>>

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their objectives and geophysical interpretations are significantly
different.
The corrections applied in eq. \ref{eq:simple-bouguer-anomaly} approximate
gravity at the geoid from measurements made on the Earth's surface by removing
the effects of the distance and the topographic masses between the two points.
Apart from the centrifugal effect mentioned in the previous section,
the gravitational effect of topographic masses between the geoid and ellipsoid
are still present in the anomaly.
Conversely, when calculating the Bouguer disturbance one removes from
the observed gravity the effects of an ellipsoidal reference model (the normal
Earth) and the known topographic masses between the surface of the ellipsoid
and observation point.
Hence, the Bouguer disturbance can be interpreted as being solely the
gravitational effect produced by the density contrasts between the actual
internal density distribution of the Earth and that of the normal Earth.

@@

\section{Gravity modelling}

We have established that the gravity disturbance is due solely to the
gravitational effect of the gravity sources.
Now we will discuss how this relates to some assumptions made during gravity
modelling.

The gravity vector $\vector{g}_i$ (eq. \ref{eq:gravity-disturbance-vector})
at a point $(x_i, y_i, z_i)$ in the TCS (Fig. \ref{fig:TCS}a) can be
represented by:

\begin{equation}
\vector{g}_i = \vector{\gamma}_i + \delta \vector{g}_i \: .
\label{eq:gravity-vector-TCS}
\end{equation}

\noindent
Furthermore, the gravity $g_i$ can be approximated by a first order Taylor's
expansion \citep{sanso_sideris2013}:

\begin{equation}
g_i \approx \gamma_i +
\versor{\gamma}_i^{\top} \delta \vector{g}_i \: ,
\label{eq:gobs-approx}
\end{equation}

\noindent
where $^{\top}$ denotes transposition and $\versor{\gamma}_{i}$ is a unit
vector with the direction of the normal gravity vector.
This approximation is valid because
$\gamma_{i} \gg \| \delta \mathbf{g}_{i} \|$ at all points on or above the
Earth's surface.
It is known in geodesy \citep[e.g.,][]{sanso_sideris2013}
and is analogous to the approximation used for
total-field magnetic anomalies \citep[e.g.,][]{blakely1996}.
In local or regional scales, the unit
vector $\versor{\gamma}_{i}$
may be considered constant throughout the study area and
parallel to the $z$ axis of the TCS (Fig. \ref{fig:TCS}a).

Using eq. \ref{eq:gobs-approx} and fixing $\gamma$ at a point $P$ at
the origin of the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance can be rewritten as:

\begin{equation}
\delta g_{i} \approx \versor{\gamma}_{P}^{\top} \delta \vector{g}_{i} \: .
\label{eq:gravity-disturbance-approx-TCS}
\end{equation}

\noindent
This equation shows that the gravity disturbance $\delta g_{i}$ is
approximately the component of the $\delta \vector{g}_{i}$ in the direction of
the normal gravity vector $\vector{\gamma}_P$
\citep{hofmann-wellenhof-moritz2005, sanso_sideris2013}.
Because $z$-axis of the TCS is parallel to $\vector{\gamma}_P$
(Fig. \ref{fig:TCS}b), the gravity disturbance can be defined as the
$z$-component of the gravitational attraction exerted by the gravity sources at
the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance can be modelled by the harmonic
function:

\begin{equation}
d_{i} = \iiint\limits_{v}
\frac{G \, \Delta\rho(x^{\prime}, y^{\prime}, z^{\prime}) \,
[z_{i} - z^{\prime}]}
{\sqrt{[x_{i} - x^{\prime}]^{2} +
[y_{i} - y^{\prime}]^{2} + [z_{i} - z^{\prime}]^{2}}} \: dv^{\prime} \: ,
\label{eq:gz-local}
\end{equation}

\noindent
where $\Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})$
is the density contrast at a point $(x^{\prime}, y^{\prime}, z^{\prime})$
within the volume $v$ and the integration is conducted over $x^{\prime}$,
$y^{\prime}$ and $z^{\prime}$.

In practice, most instances of gravity modelling in the literature use Bouguer
anomalies and function $d_{i}$ (usually presented as ``$g_z$'') to represent
the gravitational effect of gravity sources.
Consequently, these studies are using the gravity
anomaly as an approximation for the gravity disturbance.
Notice that $d_{i}$ does not depend on the
height with respect to the geoid (orthometric height).
Rather, it depends on the relative position of the gravity sources and the
observation point $(x_{i}, y_{i}, z_{i})$ in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to mention that gravimeters do not measure $d_{i}$ but
$\| \vector{g}_i \|$.
